Model Recognition as Planning

### 2018-09-20 doc/icaps-19/model-recognition.tex
paper icaps19: sections 1,2 and 3 stable. Ready for review

\begin{figure}
  \begin{scriptsize}
  \begin{center}
\begin{tikzpicture}[scale=.4]
          \begin{scope}
            \draw (0, 0) grid (5, 5);
            \node[anchor=center] at (3.5, 1.5) {R};
            \node[anchor=center] at (0.5, 0.5) {};
            \draw[thick,style=dotted] (0.5,0.5) -- (4.5,0.5);
            \draw[thick,style=dotted] (4.5,0.5) -- (4.5,1.5);
            \draw[thick,style=dotted,->] (4.5,1.5) -- (3.7,1.5);                        
          \end{scope}
        \end{tikzpicture}
  \end{center}
  \end{scriptsize}  
 \caption{\small Observation of the execution of a robot navigation plan in a $5\times 5$ grid.}
\label{fig:grid-example}
\end{figure}

To better illustrate {\em model recognition}, imagine a robot in a $n\times n$ grid whose navigation is determined by the \strips\ model of Figure~\ref{fig:model-example}. According to this model the robot could increment its {\em x-coordinate} at {\em even} rows and decrement it at the {\em odd rows}. Apart from this particular navigation model, different models could be defined within the same state variables (e.g. altering the way {\small\tt (state-q0)} and {\small\tt (state-q1)} are required and updated) and these models can determine different kinds of robot navigation. Given an observation of a plan execution, like the one illustrated at Figure~\ref{fig:grid-example}, {\em model recognition} aims here to identify which navigation model produced that observation, despite key information is unobserved (e.g. the value of {\small\tt (state-q0)} and {\small\tt (state-q1)} or the particular applied actions).

@@

\begin{figure}
  \begin{tiny}
  \begin{verbatim}
  (:action inc-x
    :parameters (?v1 ?v2)
    :precondition (and (x-coord ?v1) (next ?v1 ?v2) (state-q0))
    :effect (and (not (x-coord ?v1)) (x-coord ?v2)))

  (:action dec-x
    :parameters (?v1 ?v2)
    :precondition (and (x-coord ?v1) (next ?v2 ?v1) (state-q1))
    :effect (and (not (x-coord ?v1)) (x-coord ?v2)))

  (:action inc-y-even
    :parameters (?y1 ?y2)
    :precondition (and (y-coord ?y1) (next ?y1 ?y2) (state-q0))
    :effect (and (not (y-coord ?y1)) (y-coord ?y2)
                 (not (state-q0)) (state-q1))

  (:action inc-y-odd
    :parameters (?y1 ?y2)
    :precondition (and (y-coord ?y1) (next ?y1 ?y2) (state-q1)))
    :effect (and (not (y-coord ?y1)) (y-coord ?y2)
                 (not (state-q1)) (state-q0)))

  (:action dec-y-even
    :parameters (?y1 ?y2)
    :precondition (and (y-coord ?y1) (next ?y2 ?y1) (state-q0))
    :effect (and (not (y-coord ?y1)) (y-coord ?y2)
                 (not (state-q0)) (state-q1)))

  (:action dec-y-odd
    :parameters (?y1 ?y2)
    :precondition (and (y-coord ?y1) (next ?y2 ?y1) (state-q1))
    :effect (and (not (y-coord ?y1)) (y-coord ?y2)
                 (not (state-q1)) (state-q0)))
  \end{verbatim}           
  \end{tiny}  
 \caption{\small Action model for a robot navigation in a $n\times n$ grid.}
\label{fig:model-example}
\end{figure}

@@

For instance, Figure~\ref{fig:grid-example} illustrates a $\Phi$-observation where $\Phi$ only contains fluents of the kind {\tt\small (x-coord ?v)} and {\tt\small (y-coord ?v)}. This means that, for each observed state, only the value of fluents {\tt\small (x-coord ?v)} and {\tt\small (y-coord ?v)} is known while the value of the fluents {\tt\small (next ?v1 ?v2)}, {\tt\small (state-q0)} and {\tt\small (state-q1)} is unknown.

@@

We first formalize \strips\ schemas, then define the full space of possible \strips\ schema and eventually, we introduce an {\em edit distance} which allows us to estimate the $P(\tau|\mathcal{M})$ likelihoods for observations of the execution of classical plans and planning \strips\ models.

\begin{figure}
\begin{tiny}
\begin{verbatim}
;;; Propositional encoding for inc-x(?v1 ?v2)
(pre_x-coord_v1_inc-x) (pre_next_v1_v2_inc-x) 
(pre_state-q0__inc-x)
(del_x-coord_v1_inc-x) (add_x-coord_v2_inc-x)

;;; Propositional encoding for dec-x(?v1 ?v2)
(pre_x-coord_v1_dec-x) (pre_next_v2_v1_dec-x) 
(pre_state-q1___dec-x)
(del_x-coord_v1_dec-x) (add_x-coord_v2_dec-x)

;;; Propositional encoding for inc-y-odd(?v1 ?v2)
(pre_y-coord_v1_inc-y-odd) (pre_next_v1_v2_inc-y-odd) 
(pre_state-q0__inc-y-odd)
(del_y-coord_v1_inc-y-odd) (del_state-q1__inc-y-odd)
(add_y-coord_v2_inc-y-odd) (add_state-q0__inc-y-odd)

;;; Propositional encoding for inc-y-even(?v1 ?v2)
(pre_y-coord_v1_inc-y-even) (pre_next_v1_v2_inc-y-even)
(pre_state-q0__inc-y-even)
(del_y-coord_v1_inc-y-even) (del_state-q0__inc-y-even)
(add_y-coord_v2_inc-y-even) (add_state-q1__inc-y-even)

;;; Propositional encoding for dec-y-odd(?v1 ?v2)
(pre_y-coord_v1_dec-y-odd) (pre_next_v2_v1_dec-y-odd)
(pre_state-q1__dec-y-odd)
(del_y-coord_v1_dec-y-odd) (del_state-q1__dec-y-odd)
(add_y-coord_v2_dec-y-odd) (add_state-q0__dec-y-odd)

;;; Propositional encoding for inc-y-even(v1 ?v2)
(pre_y-coord_v1_dec-y-even) (pre_next_v2_v1_dec-y-even)
(pre_state-q0__dec-y-even)
(del_y-coord_v1_dec-y-even) (del_state-q0__dec-y-even)
(add_y-coord_v2_dec-y-even) (add_state-q1__dec-y-even)
\end{verbatim}
\end{tiny}
 \caption{\small Propositional encoding for the six schema from Figure~\ref{fig:model-example}.}
\label{fig:encoding}
\end{figure}

\subsection{A propositional encoding for \strips\ action schema}
\strips\ action schema provide a compact representation for specifying classical planning models. For instance, Figure~\ref{fig:model-example} shows six \strips\ action schema codded in PDDL for a robot navigation in a $n\times n$ grid. A \strips\ action schema $\xi$ with name, $name(\xi)$, defines a list of {\em parameters} $pars(\xi)$, and the three list of predicates ($pre(\xi)$, $del(\xi)$ and $add(\xi)$) that shape the kind of fluents that can appear in the {\em preconditions}, {\em negative effects} and {\em positive effects} of the actions induced from that schema.

We say that two \strips\ schemes $\xi$ and $\xi'$ are {\em comparable} iff $pars(\xi)=pars(\xi')$, both share the same list of parameters. For instance, we claim that the six action schema of Figure~\ref{fig:model-example} are {\em comparable} while the {\small\tt stack(?v1,?v2)} and {\small\tt pickup(?v1)} schemes from the four opertator {\em blocksworld}~\cite{slaney2001blocks} are not. Last but not least, we say that two \strips\ models $\mathcal{M}$ and $\mathcal{M}'$ are {\em comparable} iff there exists a bijective function $\mathcal{M} \mapsto \mathcal{M}^*$ that maps every action schema $\xi\in\mathcal{M}$ to a comparable schema $\xi'\in\mathcal{M'}$ and vice versa.

Given a \strips\ action schema $\xi$, a propositional encoding for the {\em preconditions}, {\em negative} and {\em positive} effects of that schema can be represented with fluents $[pre|del|add]\_p\_name(\xi)$. Figure~\ref{fig:encoding} shows the propositional encoding for the six action schema previously defined in Figure~\ref{fig:model-example}. The interest of having a propositional encoding for \strips\ action schema is that it allows to define {\em programmable actions} that is, actions whose semantics is given by the value of these particular fluents on the current state. For instance, Figure~\ref{fig:programmable} shows the programmable version of the {\tt\small inc-x(?v1,?v2)} schema for robot navigation in a $n\times n$ grid. Note that this programmable schema, when the fluents of Figure~\ref{fig:encoding} holds, behaves exactly as is defined in Figure~\ref{fig:model-example}. Further it allows to swap the semantics of two programable schemas provided that they are {\em comparable}.

\begin{figure}
  \begin{tiny}  
  \begin{verbatim}
(:action inc-x
  :parameters (?v1 ?v2)
  :precondition
    (and (or (not (pre_x-coord_v1_inc-x)) (x-coord ?v1))
         (or (not (pre_x-coord_v2_inc-x)) (x-coord ?v2))
         (or (not (pre_y-coord_v1_inc-x)) (x-coord ?v1))                       
         (or (not (pre_y-coord_v2_inc-x)) (x-coord ?v2))
         (or (not (pre_state-q0__inc-x)) (state-q0))
         (or (not (pre_state-q1__inc-x)) (state-q1)))
         (or (not (pre_next_v1_v1_inc-x)) (next ?v1 ?v1)))
         (or (not (pre_next_v1_v2_inc-x)) (next ?v1 ?v2)))
         (or (not (pre_next_v2_v1_inc-x)) (next ?v2 ?v1)))
         (or (not (pre_next_v2_v2_inc-x)) (next ?v2 ?v2)))         
    )
    :effect (and
       (when (del_x-coord_v1_inc-x) (not (x-coord ?v1)))
       (when (del_x-coord_v2_inc-x) (not (x-coord ?v2)))
       (when (del_y-coord_v1_inc-x) (not (x-coord ?v1)))
       (when (del_y-coord_v2_inc-x) (not (x-coord ?v2)))
       (when (del_state-q0__inc-x) (not (state-q0)))
       (when (del_state-q1__inc-x) (not (state-q1)))
       (when (del_next_v1_v1_inc-x) (not (next ?v1 ?v1)))
       (when (del_next_v1_v2_inc-x) (not (next ?v1 ?v2)))
       (when (del_next_v2_v1_inc-x) (not (next ?v2 ?v1)))
       (when (del_next_v2_v2_inc-x) (not (next ?v2 ?v2)))

       (when (add_x-coord_v1_inc-x) (x-coord ?v1))
       (when (add_x-coord_v2_inc-x) (x-coord ?v2))
       (when (add_y-coord_v1_inc-x) (x-coord ?v1))
       (when (add_y-coord_v2_inc-x) (x-coord ?v2))
       (when (add_state-q0__inc-x) (state-q0))
       (when (add_state-q1__inc-x) (state-q1))
       (when (add_next_v1_v1_inc-x) (next ?v1 ?v1))
       (when (add_next_v1_v2_inc-x) (next ?v1 ?v2))
       (when (add_next_v2_v1_inc-x) (next ?v2 ?v1))
       (when (add_next_v2_v2_inc-x) (next ?v2 ?v2)))
  \end{verbatim}           
  \end{tiny}  
 \caption{\small Programmable version of the {\tt\small inc-x(?v1,?v2)} schema for robot navigation in a $n\times n$ grid.}
\label{fig:programmable}
\end{figure}

\subsection{The space of \strips\ schema}
The space of possible \strips\ schema is bound by these elements:
\begin{enumerate}
\item $\Psi$, is the set of {\em state predicates} that shape the propositional state variables of the model.
\item $pars(\xi)$, the list of {\em parameters} of the corresponding schema $\xi$. 
\end{enumerate}

For each action schema $\xi$, given the set of {\em state predicates} $\Psi$, then $F_{\xi}$ defines the subset of elements that can appear in the preconditions and effects of that action schema. For the actions schema defined in~\ref{fig:model-example} this set is the same and contains the following ten elements, {\small\tt\{x-coord($v_1$), x-coord($v_2$), y-coord($v_1$), y-coord($v_2$), state-q0(), state-q1(), next($v_1$,$v_1$), next($v_1$,$v_2$), next($v_2$,$v_1$), next($v_2$,$v_2$))\}}.

\begin{enumerate}
\setcounter{enumi}{2}
\item ${\mathcal C}$, a set of {\em syntactic constraints}. These constraints include the \strips\ constraints (we assume that $\eff^-(a)\subseteq \pre(a)$, $\eff^-(a)\cap \eff^+(a)=\emptyset$ and $\pre(a)\cap \eff^+(a)=\emptyset$) and domain-specific constraints. For instace, in a {\em robot navigation} domain like the modeled in Figure~\ref{fig:model-example}, predicates {\small\tt state-q0()} and {\small\tt state-q1()} are exclusive so they cannot hold a the same time.
\end{enumerate}

Predicates $\Psi_a$ also shape an additional set of objects ($\Omega\cap\Omega_v=\emptyset$) that we denote as {\em variable names} and that is bound by the maximum arity of the given action headers, i.e., $\Omega_v=\{v_i\}_{i=1}^{\operatorname*{max}_{\xi} ar(\xi)}$. For instance, in a $5\times 5$ grid (like the one in Figure~\label{fig:grid-example} and modeled in Figure~\ref{fig:model-example}) $\Omega=\{o_1, o_2, o_3, o_4, o_5\}$ while $\Omega_v=\{v_1, v_2\}$ because all the actions schema have arity two.

\begin{definition}[The space of \strips\ models]
Given the set of {\em state predicates} $\Psi$, the set of {\em variable names} $\Omega_v$ and the set of {\em syntactic constraints} ${\mathcal C}$, the space of models for a given op predicate $p_\xi$ that represents the header of an action schema $\xi$ is given by three sets of inter
\end{definition}

The size of the space of possible \strips\ models for an action schema $\xi$ is given by the expression, $2^{2\times|F_{\xi}|}$ (\strips\ constraints require negative effects appearing as preconditions, negative effects cannot be positive effects at the same time and also, positive effects cannot appear as preconditions). For the mentioned navigation model, $2^{2\times|F_{\xi}|}=1,048,576$ for every action schema.

@@

\section{Model recognition as planning}
This section shows that, when the dynamics of the actions $A\in P$ are specified with \strips\ action schemas, then $\delta(\mathcal{M},\tau)$ can be computed with a compilation of a classical planning with conditional effects. The intuition behind this compilation is that a solution to the resulting classical planning task is a sequence of actions that:
\begin{enumerate}
\item {\bf Edits the action model $\mathcal{M}$ to build $\mathcal{M}'$}. A solution plan starts with a {\em prefix} that modifies the preconditions and effects of the action schemes in $\mathcal{M}$ using to the two {\em edit operations} defined above, {\em deletion} and {\em insertion}. 
\item {\bf Validates the edited model $\mathcal{M}'$}. The solution plan continues with a postfix that:
\begin{enumerate}
\item Induces an optimal solution plan $\pi^*_\tau$ for the original classical planning problem $P$.
\item Validates that $\tau$ is an observation of the execution of $\pi^*_\tau$ on the classical planning problem $P$.
\end{enumerate}
\end{enumerate}

@@

@@

\item Actions for {\em applying} an edited operator schema $\xi\in\mathcal{M}$ bound with objects $\omega\subseteq\Omega^{ar(\xi)}$. Since operators headers are given as input, the variables $pars(\xi)$ are bound to the objects in $\omega$ that appear at the same position. Figure~\ref{fig:programmable} shows the PDDL encoding of the action for applying a programmed operator.
\begin{small}
\begin{align*}
\hspace*{7pt}\pre(\mathsf{apply_{\xi,\omega}})=&\{pre_{f}(\xi)\implies p(\omega)\}_{\forall p\in\Psi,f=p(pars(\xi))}\\
&\cup \{\neg mode_{val}\},\\
\cond(\mathsf{apply_{\xi,\omega}})=&\{del_{f}(\xi)\}\rhd\{\neg p(\omega)\}_{\forall p\in\Psi,f=p(pars(\xi))},\\
&\{add_{f}(\xi)\}\rhd\{p(\omega)\}_{\forall p\in\Psi,f=p(pars(\xi))},\\
&\{mode_{edit}\}\rhd\{\neg mode_{edit}\},\\
&\{\emptyset\}\rhd\{mode_{val}\}.
\end{align*}
\end{small}

@@

\subsection{Recognition of {\em Regular Automatae}}
We analyze now {\em model recognition} when the input set of given set of models represent different {\em regular automatae}.

A {\em Regular automatae} is a tuple $\mathcal{M}=\tup{Q,q_o,Q_{\bot},\Sigma\square,\delta}$:
\begin{itemize}
\item $Q$ is a finite and non-empty set of machine states with the {\em initial state} $q_0\in Q$ and the {\em terminal states} $Q_{\bot}\subseteq Q$.  
\item $\Sigma$ is the {\em input alphabet}, a finite non-empty set of symbols and the {\em blank symbol} $\square\in\Upsilon$ (the only symbol allowed to occur on the tape infinitely often).
\item $\delta: \Sigma\times (Q\setminus Q_{\bot}) \rightarrow \Sigma\times Q\times\{left,right\}$ is the {\em transition function}. For each pair of tape symbol and non-terminal machine state $\delta$ defines: (1), the tape symbol to print at the current position of the header (2), the new state of the machine and (3), whether the header is shifted {\em left} or {\em right} after the print operation. If $\delta$ is not defined for the current pair of tape symbol and machine state, the machine halts.
\end{itemize}

>>>>

\begin{figure}
  \begin{scriptsize}    
  \begin{center}
  \begin{tikzpicture}[scale=.4]
          \begin{scope}
            \draw (0, 0) grid (5, 5);
            \node[anchor=center] at (3.5, 1.5) {R};
            \node[anchor=center] at (0.5, 0.5) {};
            \draw[thick,style=dotted] (0.5,0.5) -- (4.5,0.5);
            \draw[thick,style=dotted] (4.5,0.5) -- (4.5,1.5);
            \draw[thick,style=dotted,->] (4.5,1.5) -- (3.7,1.5);                        
          \end{scope}
        \end{tikzpicture}
\hspace*{1cm}        
	\begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=2cm,semithick]
	  \node[state] (A)              {$q_0$};
	  \node[state] (B) [right of=A] {$q_1$};          
	  \path
(A) edge [bend left] node {inc-y, dec-y} (B)
(B) edge [bend left] node {inc-y, dec-y} (A);
	\end{tikzpicture}
  \end{center}
  \end{scriptsize}  
 \caption{\small Observation of the execution of a robot navigation plan in a $5\times 5$ grid and a two state automata for controlling that the robot only increments its x-coordinate at {\em even} rows (i.e. at state $q_0$).}
\label{fig:grid-example}
\end{figure}

To better illustrate {\em model recognition}, imagine a robot in a $n\times n$ grid whose navigation is determined by the \strips\ model of Figure~\ref{fig:model-example}. According to this model the robot could increment its {\em x-coordinate} at {\em even} rows and decrement it at the {\em odd rows}. Apart from this particular navigation model, different models could be defined within the same state variables (e.g. altering the way {\small\tt (q0)} and {\small\tt (q1)} are required and updated) and these models can determine different kinds of robot navigation. Given an observation of a plan execution, like the one illustrated at Figure~\ref{fig:grid-example}, {\em model recognition} aims here to identify which navigation model produced that observation, despite key information is unobserved (e.g. the value of {\small\tt (q0)} and {\small\tt (q1)} or the particular applied actions).

@@

\begin{figure}
  \begin{tiny}
  \begin{verbatim}
  (:action inc-x
    :parameters (?v1 ?v2)
    :precondition (and (xcoord ?v1) (next ?v1 ?v2) (q0))
    :effect (and (not (xcoord ?v1)) (xcoord ?v2)))

  (:action dec-x
    :parameters (?v1 ?v2)
    :precondition (and (xcoord ?v1) (next ?v2 ?v1) (q1))
    :effect (and (not (xcoord ?v1)) (xcoord ?v2)))

  (:action inc-y-even
    :parameters (?y1 ?y2)
    :precondition (and (ycoord ?y1) (next ?y1 ?y2) (q0))
    :effect (and (not (ycoord ?y1)) (ycoord ?y2)
                 (not (q0)) (q1))

  (:action inc-y-odd
    :parameters (?y1 ?y2)
    :precondition (and (ycoord ?y1) (next ?y1 ?y2) (q1)))
    :effect (and (not (ycoord ?y1)) (ycoord ?y2)
                 (not (q1)) (q0)))

  (:action dec-y-even
    :parameters (?y1 ?y2)
    :precondition (and (ycoord ?y1) (next ?y2 ?y1) (q0))
    :effect (and (not (ycoord ?y1)) (ycoord ?y2)
                 (not (q0)) (q1)))

  (:action dec-y-odd
    :parameters (?y1 ?y2)
    :precondition (and (ycoord ?y1) (next ?y2 ?y1) (q1))
    :effect (and (not (ycoord ?y1)) (ycoord ?y2)
                 (not (q1)) (q0)))
  \end{verbatim}           
  \end{tiny}  
 \caption{\small Action model for a robot navigation in a $n\times n$ grid.}
\label{fig:model-example}
\end{figure}

@@

For instance, Figure~\ref{fig:grid-example} illustrates the $\Phi$-observation \{{\tt\scriptsize<(xcoord 1)(ycoord 1)>, <(xcoord 2)(ycoord 1)>, <(xcoord 3)(ycoord 1)>, <(xcoord 4)(ycoord 1)>, <(xcoord 5)(ycoord 1)>, <(xcoord 5)(ycoord 2)>, <(xcoord 4)(ycoord 2)>}\} where $\Phi$ only contains fluents of the kind {\tt\small (xcoord ?v)} and {\tt\small (ycoord ?v)}. This means that, for each observed state, only the value of fluents {\tt\small (xcoord ?v)} and {\tt\small (ycoord ?v)} is known while the value of the fluents {\tt\small (next ?v1 ?v2)}, {\tt\small (q0)} and {\tt\small (q1)} is unknown.

@@

We first formalize \strips\ schemas and define the full space of possible \strips\ schema. Eventually, we introduce an {\em edit distance} which allows us to estimate the $P(\tau|\mathcal{M})$ likelihoods for observations of the execution of classical plans and planning \strips\ models.

\subsection{The space of \strips\ schema}
\strips\ action schema provide a compact representation for specifying classical planning models. For instance, Figure~\ref{fig:model-example} shows six \strips\ action schema codded in PDDL for a robot navigation in a $n\times n$ grid.

\begin{definition}[A \strips\ action schema $\xi$]
A \strips\ action schema $\xi$ with name, $name(\xi)$, defines a list of {\em parameters} $pars(\xi)$, and the three list of predicates ($pre(\xi)$, $del(\xi)$ and $add(\xi)$) that shape the kind of fluents that can appear in the {\em preconditions}, {\em negative effects} and {\em positive effects} of the actions induced from that schema. 
\end{definition}

Let us define additional set of objects ($\Omega\cap\Omega_\xi=\emptyset$) that we denote as {\em variable names} and that contains one variable name for each paramter in $pars(\xi)$, that is $\Omega_\xi=\{v_i\}_{i=1}^{|pars(\xi)|}$. For instance, for all the \strips\ schema shown in Figure~\ref{fig:model-example} $\Omega_\xi=\{v_1,v_2\}$. Note that, for any of the actions schema defined in~\ref{fig:model-example}, the FOL interpretations of $\Psi$ (the {\em state predicates} that shape the propositional state variables) over the corresponding set of objects $\Omega_\xi$ contains the same ten elements, {\small\tt\{xcoord($v_1$), xcoord($v_2$), ycoord($v_1$), ycoord($v_2$), q0(), q1(), next($v_1$,$v_1$), next($v_1$,$v_2$), next($v_2$,$v_1$), next($v_2$,$v_2$))\}}.

In addition, we can define constraints that bound the space of possible \strips\ schema:
${\mathcal C}$
\begin{itemize}
\item {\em Syntactic constraints}. These constraints include the \strips\ constraints (we assume that $\eff^-(a)\subseteq \pre(a)$, $\eff^-(a)\cap \eff^+(a)=\emptyset$ and $\pre(a)\cap \eff^+(a)=\emptyset$). The size of the space of possible \strips\ schema is given by the expression, $2^{2\times|F_{\xi}|}$ (\strips\ constraints require negative effects appearing as preconditions, negative effects cannot be positive effects at the same time and also, positive effects cannot appear as preconditions). For the mentioned navigation model, $2^{2\times|F_{\xi}|}=1,048,576$ for every action schema.
\item {\em Domain-specific constraints}. For instace, in a {\em robot navigation} model, like the one in Figure~\ref{fig:model-example}, predicates {\small\tt q0()} and {\small\tt q1()} are exclusive so they cannot hold a the same time. Further, next($v_1$,$v_1$) and next($v_2$,$v_2$) is not appearing neither at the preconditions nor the effects of any possible \strips\ schema. This domain-specific knowledge can be used to constraint further the space of possible \strips\ schema.
\end{itemize}

Now we are ready to define the space of \strips\ schema that are possible for $\xi$.
\begin{definition}[The space of possible \strips\ schema for $\xi$]
The space of possible \strips\ schema for $\xi$ is the set ${\mathcal I}_{\Psi,\xi,{\mathcal C}}$ of the FOL interpretations of $\Psi$ (the {\em state predicates} that shape the propositional state variables) over the corresponding set of objects $\Omega_\xi$ and that satisfy the given shcema constraints.
\end{definition}

@@

\section{Model recognition with classical planning}
This section shows that, when the dynamics of the actions $A\in P$ are specified with \strips\ action schemas, then $\delta(\mathcal{M},\tau)$ can be computed with a compilation of a classical planning with conditional effects. The intuition behind this compilation is that a solution to the resulting classical planning task is a sequence of actions that:
\begin{enumerate}
\item {\bf Edits the action model $\mathcal{M}$ to build $\mathcal{M}'$}. A solution plan starts with a {\em prefix} that modifies the preconditions and effects of the action schemes in $\mathcal{M}$ using to the two {\em edit operations} defined above, {\em deletion} and {\em insertion}. 
\item {\bf Validates the edited model $\mathcal{M}'$}. The solution plan continues with a postfix that:
\begin{enumerate}
\item Induces an optimal solution plan $\pi^*_\tau$ for the original classical planning problem $P$.
\item Validates that $\tau$ is an observation of the execution of $\pi^*_\tau$ on the classical planning problem $P$.
\end{enumerate}
\end{enumerate}

@@

\begin{figure}
\begin{tiny}
\begin{verbatim}
;;; Propositional encoding for inc-x(?v1 ?v2)
(pre_xcoord_v1_inc-x) (pre_next_v1_v2_inc-x) 
(pre_q0__inc-x)
(del_xcoord_v1_inc-x) (add_xcoord_v2_inc-x)

;;; Propositional encoding for dec-x(?v1 ?v2)
(pre_xcoord_v1_dec-x) (pre_next_v2_v1_dec-x) 
(pre_q1___dec-x)
(del_xcoord_v1_dec-x) (add_xcoord_v2_dec-x)

;;; Propositional encoding for inc-y-even(?v1 ?v2)
(pre_ycoord_v1_inc-y-even) (pre_next_v1_v2_inc-y-even)
(pre_q0__inc-y-even)
(del_ycoord_v1_inc-y-even) (del_q0__inc-y-even)
(add_ycoord_v2_inc-y-even) (add_q1__inc-y-even)

;;; Propositional encoding for inc-y-odd(?v1 ?v2)
(pre_ycoord_v1_inc-y-odd) (pre_next_v1_v2_inc-y-odd) 
(pre_q0__inc-y-odd)
(del_ycoord_v1_inc-y-odd) (del_q1__inc-y-odd)
(add_ycoord_v2_inc-y-odd) (add_q0__inc-y-odd)

;;; Propositional encoding for dec-y-even(v1 ?v2)
(pre_ycoord_v1_dec-y-even) (pre_next_v2_v1_dec-y-even)
(pre_q0__dec-y-even)
(del_ycoord_v1_dec-y-even) (del_q0__dec-y-even)
(add_ycoord_v2_dec-y-even) (add_q1__dec-y-even)

;;; Propositional encoding for dec-y-odd(?v1 ?v2)
(pre_ycoord_v1_dec-y-odd) (pre_next_v2_v1_dec-y-odd)
(pre_q1__dec-y-odd)
(del_ycoord_v1_dec-y-odd) (del_q1__dec-y-odd)
(add_ycoord_v2_dec-y-odd) (add_q0__dec-y-odd)
\end{verbatim}
\end{tiny}
 \caption{\small Propositional encoding for the six schema from Figure~\ref{fig:model-example}.}
\label{fig:encoding}
\end{figure}

\subsection{A propositional encoding for \strips\ action schema}
We say that two \strips\ schemes $\xi$ and $\xi'$ are {\em comparable} iff $pars(\xi)=pars(\xi')$, both share the same list of parameters. For instance, we claim that the six action schema of Figure~\ref{fig:model-example} are {\em comparable} while the {\small\tt stack(?v1,?v2)} and {\small\tt pickup(?v1)} schemes from the four opertator {\em blocksworld}~\cite{slaney2001blocks} are not. Last but not least, we say that two \strips\ models $\mathcal{M}$ and $\mathcal{M}'$ are {\em comparable} iff there exists a bijective function $\mathcal{M} \mapsto \mathcal{M}^*$ that maps every action schema $\xi\in\mathcal{M}$ to a comparable schema $\xi'\in\mathcal{M'}$ and vice versa.

Given a \strips\ action schema $\xi$, a propositional encoding for the {\em preconditions}, {\em negative} and {\em positive} effects of that schema can be represented with fluents $[pre|del|add]\_p\_name(\xi)$. Figure~\ref{fig:encoding} shows the propositional encoding for the six action schema previously defined in Figure~\ref{fig:model-example}. The interest of having a propositional encoding for \strips\ action schema is that it allows to define {\em editable actions} that is, actions whose semantics is given by the value of these particular fluents on the current state.

Formally given an operator schema $\xi\in\mathcal{M}$ we can build its {\em editable} version as follows:
\begin{small}
\begin{align*}
\hspace*{7pt}\pre(\mathsf{editable_{\xi,\omega}})=&\{pre_{f}(\xi)\implies p(\omega)\}_{\forall p\in\Psi,f=p(pars(\xi))}\\
\cond(\mathsf{editable_{\xi,\omega}})=&\{del_{f}(\xi)\}\rhd\{\neg p(\omega)\}_{\forall p\in\Psi,f=p(pars(\xi))},\\
&\{add_{f}(\xi)\}\rhd\{p(\omega)\}_{\forall p\in\Psi,f=p(pars(\xi))}.
\end{align*}
\end{small}
bound with objects $\omega\subseteq\Omega^{ar(\xi)}$. Since operators headers are given as input, the variables $pars(\xi)$ are bound to the objects in $\omega$ that appear at the same position. Figure~\ref{fig:editable} shows the PDDL encoding of the action for applying a programmed operator. For instance, Figure~\ref{fig:editable} shows the editable version of the {\tt\small inc-x(?v1,?v2)} schema for robot navigation in a $n\times n$ grid (see Figure~\ref{fig:model-example}). Note that this editable schema, when the fluents of Figure~\ref{fig:encoding} holds, behaves exactly as is defined in Figure~\ref{fig:model-example}. Further it allows to swap the semantics of two programable schemas provided that they are {\em comparable}.

\begin{figure}
  \begin{tiny}  
  \begin{verbatim}
(:action inc-x
  :parameters (?v1 ?v2)
  :precondition
    (and (or (not (pre_xcoord_v1_inc-x)) (xcoord ?v1))
         (or (not (pre_xcoord_v2_inc-x)) (xcoord ?v2))
         (or (not (pre_ycoord_v1_inc-x)) (xcoord ?v1))                       
         (or (not (pre_ycoord_v2_inc-x)) (xcoord ?v2))
         (or (not (pre_q0__inc-x)) (q0))
         (or (not (pre_q1__inc-x)) (q1)))
         (or (not (pre_next_v1_v1_inc-x)) (next ?v1 ?v1)))
         (or (not (pre_next_v1_v2_inc-x)) (next ?v1 ?v2)))
         (or (not (pre_next_v2_v1_inc-x)) (next ?v2 ?v1)))
         (or (not (pre_next_v2_v2_inc-x)) (next ?v2 ?v2))))
    :effect (and
       (when (del_xcoord_v1_inc-x) (not (xcoord ?v1)))
       (when (del_xcoord_v2_inc-x) (not (xcoord ?v2)))
       (when (del_ycoord_v1_inc-x) (not (xcoord ?v1)))
       (when (del_ycoord_v2_inc-x) (not (xcoord ?v2)))
       (when (del_q0__inc-x) (not (q0)))
       (when (del_q1__inc-x) (not (q1)))
       (when (del_next_v1_v1_inc-x) (not (next ?v1 ?v1)))
       (when (del_next_v1_v2_inc-x) (not (next ?v1 ?v2)))
       (when (del_next_v2_v1_inc-x) (not (next ?v2 ?v1)))
       (when (del_next_v2_v2_inc-x) (not (next ?v2 ?v2)))

       (when (add_xcoord_v1_inc-x) (xcoord ?v1))
       (when (add_xcoord_v2_inc-x) (xcoord ?v2))
       (when (add_ycoord_v1_inc-x) (xcoord ?v1))
       (when (add_ycoord_v2_inc-x) (xcoord ?v2))
       (when (add_q0__inc-x) (q0))
       (when (add_q1__inc-x) (q1))
       (when (add_next_v1_v1_inc-x) (next ?v1 ?v1))
       (when (add_next_v1_v2_inc-x) (next ?v1 ?v2))
       (when (add_next_v2_v1_inc-x) (next ?v2 ?v1))
       (when (add_next_v2_v2_inc-x) (next ?v2 ?v2)))
  \end{verbatim}           
  \end{tiny}  
 \caption{\small Editable version of the {\tt\small inc-x(?v1,?v2)} schema for robot navigation in a $n\times n$ grid.}
\label{fig:editable}
\end{figure}

@@

\item Actions for {\em applying} an edited operator schema $\xi\in\mathcal{M}$ bound with objects $\omega\subseteq\Omega^{ar(\xi)}$. Since operators headers are given as input, the variables $pars(\xi)$ are bound to the objects in $\omega$ that appear at the same position. Figure~\ref{fig:editable} shows the PDDL encoding of the action for applying a programmed operator.
\begin{small}
\begin{align*}
\hspace*{7pt}\pre(\mathsf{apply_{\xi,\omega}})=&\{pre_{f}(\xi)\implies p(\omega)\}_{\forall p\in\Psi,f=p(pars(\xi))}\\
&\cup \{\neg mode_{val}\},\\
\cond(\mathsf{apply_{\xi,\omega}})=&\{del_{f}(\xi)\}\rhd\{\neg p(\omega)\}_{\forall p\in\Psi,f=p(pars(\xi))},\\
&\{add_{f}(\xi)\}\rhd\{p(\omega)\}_{\forall p\in\Psi,f=p(pars(\xi))},\\
&\{mode_{edit}\}\rhd\{\neg mode_{edit}\},\\
&\{\emptyset\}\rhd\{mode_{val}\}.
\end{align*}
\end{small}

@@

\subsection{Recognition of {\em Regular Automatae}}
We analyze now {\em model recognition} when the given set of models represent different {\em regular automatae}. A {\em Regular automatae} is a tuple $\mathcal{M}=\tup{Q,q_o,Q_{\bot},\Sigma\square,\delta}$:
\begin{itemize}
\item $Q$ is a finite and non-empty set of machine states with the {\em initial state} $q_0\in Q$ and the {\em terminal states} $Q_{\bot}\subseteq Q$.  
\item $\Sigma$ is the {\em input alphabet}, a finite non-empty set of symbols and the {\em blank symbol} $\square\in\Upsilon$ (the only symbol allowed to occur on the tape infinitely often).
\item $\delta: \Sigma\times (Q\setminus Q_{\bot}) \rightarrow \Sigma\times Q\times\{left,right\}$ is the {\em transition function}. For each pair of tape symbol and non-terminal machine state $\delta$ defines: (1), the tape symbol to print at the current position of the header (2), the new state of the machine and (3), whether the header is shifted {\em left} or {\em right} after the print operation. If $\delta$ is not defined for the current pair of tape symbol and machine state, the machine halts.
\end{itemize}
